---
title: "taggart"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE
)
```

This package is a wrapper for downloading tagging data for mackerel using the API of [Norwegian Marine Data Centre](https://nmdc.no/nmdc). Any use of the data on mackerel should give reference to [Slotte et al (2021)](#ref1). The package is planned to be extended with tagging data for Norwegian spring spawning herring, but this is a work in progress.



The taggart package was originally developed by Einar Hjorleifsson, but Sondre Hølleland has taken over the maintenance and further development of the package. 

## Authors' github accounts

**Einar Hjorleifsson** - [einarhjorleifsson](https://github.com/einarhjorleifsson)

**Sondre Hølleland** - [holleland](https://github.com/holleland)


## <a name="ref1"></a>Data reference

Aril Slotte (IMR), Anna Ólafsdóttir (MFRI), Sigurður Þór Jónsson (MFRI), Jan Arge Jacobsen (FAMRI) and Steve Mackinson (SPFA) (2021) PIT-tag time series for studying migrations and use in stock assessment of North East Atlantic mackerel (Scomber Scombrus) http://metadata.nmdc.no/metadata-api/landingpage/f9e8b1cff4261cf6575e70e56c4c3b3e



## License

This project is licensed under the GNU GPLv3 License - see [LICENSE](LICENSE.md). The data is licensed under Creative Commons [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/) license (see [Slotte et al (2021)](#ref1)).

Please contact leader of this tagging program, Aril Slotte <aril@hi.no> prior to use of data to secure a full understanding of the data, to avoid any conflicts of interest with ongoing research based on these data and to discuss potential research cooperation.

## Example code

To install the package, run the following code:
```{r, eval = FALSE}
devtools::install_github("IMRpelagic/taggart", dependencies = FALSE)
```

Load library:
```{r, message = FALSE}
library(taggart)
```

Get the data and have a peek
```{r}
tg_catches() %>% glimpse()
tg_catches_bio() %>% glimpse()
tg_catches_link() %>% glimpse()
tg_expeditions() %>% glimpse()
tg_expeditions_bio() %>% glimpse()
tg_expeditions_link() %>% glimpse()
tg_outoforder() %>% glimpse()
```

## Plot map of recaptures
As an example, we can download the catch data using the API and create a plot of the catches by year.
```{r, echo = TRUE, fig.width = 8, fig.height = 4, message = FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
# Download the data
d <- tg_catches()

# Calculate range of longitudes and latitudes
xlim = range(d$cLon)
ylim = range(d$cLat)
xlim = quantile(d$cLon, probs = c(0.01, 0.99))
ylim = quantile(d$cLat, probs = c(0.01, 0.99))

# Get map: 
m <- ggplot2::map_data("world",
                       xlim = xlim * c(0.9, 1.1),
                       ylim = ylim * c(0.9, 1.1))
# Create plot: 
d %>% 
  mutate(year = year(CatchDate)) %>% 
  group_by(year, cLon, cLat) %>% 
  summarise(wgt = sum(CatchWeight) / 1e6,
            .groups = "drop") %>% 
  mutate(wgt = case_when(wgt <  2 ~ "00-02",
                         wgt <  5 ~ "02-05",
                         wgt < 10 ~ "05-10",
                         wgt < 25 ~ "10-25",
                         TRUE ~ "25+")) %>% 
  ggplot() +
  theme_void() +
  geom_polygon(data = m, aes(long, lat, group = group), fill = "grey") +
  geom_tile(aes(cLon, cLat, fill = wgt)) +
  coord_quickmap(xlim = xlim, ylim = ylim) +
  facet_wrap( ~year, ncol = 5) +
  guides(color = guide_colorbar(barheight = 15))+
  scale_fill_viridis_d(option = "B", direction = -1)
```

